---
title: "Generate figures"
output: html_notebook
---

# Setup
```{r}
library(irr)
library(ggplot2)

workdir <- '/path/to/workdir'

```



# Figure 4: Fitted and imputed enCCA canonical correlations across subsamples
```{r}
canoncorrs1 <- read.csv(paste(workdir, '/replication/results/elasticnetCCA_scaled/allsubsets_canoncorrs1.csv', sep = ""), header = FALSE)
names(canoncorrs1) <- c("corr", "pval", "CI95lower", "CI95upper")
canoncorrs1$CV <- 1
canoncorrs1$Subsample <- c("Discovery", "Test", "Replication")

canoncorrs2 <- read.csv(paste(workdir, '/replication/results/elasticnetCCA_scaled/allsubsets_canoncorrs2.csv', sep = ""), header = FALSE)
names(canoncorrs2) <- c("corr", "pval", "CI95lower", "CI95upper")
canoncorrs2$CV <- 2
canoncorrs2$Subsample <- c("Discovery", "Test", "Replication")

canoncorrs3 <- read.csv(paste(workdir, '/replication/results/elasticnetCCA_scaled/allsubsets_canoncorrs3.csv', sep = ""), header = FALSE)
names(canoncorrs3) <- c("corr", "pval", "CI95lower", "CI95upper")
canoncorrs3$CV <- 3
canoncorrs3$Subsample <- c("Discovery", "Test", "Replication")

figdata <- rbind(canoncorrs1, canoncorrs2, canoncorrs3)
figdata[1,3] <- NA
figdata[1,4] <- NA
figdata[4,3] <- NA
figdata[4,4] <- NA
figdata[7,3] <- NA
figdata[7,4] <- NA

ggplot(figdata, aes(x=CV, y=corr, fill=Subsample, color=Subsample)) +
  geom_bar(stat="identity", position = "dodge", alpha=0.7) +
  scale_fill_manual(values = c("#48AEE1", "#EC6C53","#67C60F")) +
  scale_color_manual(values = c("#48AEE1", "#EC6C53","#67C60F")) +
  geom_errorbar( aes(x=CV, ymin=CI95lower, ymax=CI95upper), alpha=0.9, size=0.5, position = "dodge") +
  labs(x="Canonical variate pair (signature)", y="Canonical correlation") +
  theme_bw() + 
  theme(panel.grid.minor = element_blank(), panel.grid.major.x = element_blank()) +
  theme(legend.position = c(0.9, 0.85)) +
  theme(panel.border = element_blank(), axis.line = element_line()) +
  geom_hline(yintercept=0, linetype="dashed")


ggsave(paste(workdir, '/figures/fig4.png', sep = ""), width = 8, height = 6, dpi = 1000)
```


# Figure xx: enCCA feature loadings

## EnCCA symptom loadings

```{r}
# Libraries
library(ggplot2)
library(svglite)

# CCA clinical variable loadings
signflip <- c(-1, -1, -1, -1, 1, -1, -1, 1)
clin_labels <- c("Obsessions", "SomethingBad", "Perfect", "Guilty", "Compulsions", "StrangeBehavior", "StrangeIdeas", "Worries")

i <- 1
for (i in 1:length(clin_labels)) {
  print(i)
  clin_loadings <- read.table(paste(workdir, '/replication/results/elasticnetCCA_scaled/replication_selected_features/allreplication_clinloadings_u_CV', i, 'norm.csv', sep = ""))
  clin_loadings <- as.data.frame(clin_loadings * signflip[i])
  names(clin_loadings) <- "Loadings"
  clin_loadings$Symptoms <- factor(clin_labels, levels = rev(clin_labels), ordered = TRUE)
  levels(clin_loadings$Symptoms)
  
  if(sum(clin_loadings$Loadings>0) == 8) { #if all loadings are positive
    clin_visual <- ggplot(data=clin_loadings, aes(x=Symptoms, y=Loadings, fill=Loadings>0)) + 
    geom_bar(stat="identity") + 
    scale_fill_manual(values = c("blue")) +
    #discovery
    #ylim(-2500,3140) +
    #replication
    ylim(-50,60) +
    #geom_text(aes(label = Symptoms, vjust = 0)) +
    labs(x = "", y = "Loading") +
    coord_flip() +
    theme_classic() +
    theme(axis.text = element_text(size = 17), axis.title = element_text(size = 17)) +
    theme (legend.position = "none")
  } else {
    clin_visual <- ggplot(data=clin_loadings, aes(x=Symptoms, y=Loadings, fill=Loadings>0)) + 
      geom_bar(stat="identity") + 
      scale_fill_manual(values = c("red", "blue")) +
      #discovery
      #ylim(-2500,3140) +
      #replication
      ylim(-50,60) +
      #geom_text(aes(label = Symptoms, vjust = 0)) +
      labs(x = "", y = "Loading") +
      coord_flip() +
      theme_classic() +
      theme(axis.text = element_text(size = 17), axis.title = element_text(size = 17)) +
      theme (legend.position = "none")
  }
  
  clin_visual
  
  #save plot
  png(paste('/Users/tracey/Desktop/OCS_bigdata_ML/results/replication/comparison/plot_allreplication_clinloadings_CV', i, '.png', sep = ""))
  clin_visual
  dev.off() 
  
}
```

```{r}
clin_loadings_disc <- read.table(paste(workdir, '/discovery/results/elasticnetCCA_scaled/alldiscovery_clinloadings_u_CV1norm.csv', sep = ""))
clin_loadings_replA <- read.table(paste(workdir, '/replication/results/elasticnetCCA_scaled/discovery-selected-features/allreplication_clinloadings_u_CV1norm.csv', sep = ""))
bothclinloadingsA <- cbind(clin_loadings_disc, clin_loadings_replA)
icc(bothclinloadingsA, model = "twoway", unit = "average", type = "consistency")

clin_loadings_replB <- read.table(paste(workdir, '/replication/results/elasticnetCCA_scaled/allreplication_clinloadings_u_CV1norm.csv', sep = ""))
bothclinloadingsB <- cbind(clin_loadings_disc, clin_loadings_replB)
icc(bothclinloadingsB, model = "twoway", unit = "average", type = "consistency")
```


## EnCCA rsFC loadings
Figures made in Brain Net Viewer in Matlab

## Compare rsFC loadings across subsamples (Discovery versus Replication with Discovery-selected features)
```{r}
rsfc_loadings_disc <- read.table(paste(workdir, '/discovery/results/elasticnetCCA_scaled/alldiscovery_rsFCloadings_v_CV1.csv', sep = ""))
rsfc_inds_disc <- read.table(paste(workdir, '/discovery/results/elasticnetCCA_scaled/alldiscovery_rsFCinds_v_CV1.csv', sep = ""))
use_rsfc_loadings_disc <- cbind(unlist(rsfc_inds_disc),unlist(rsfc_loadings_disc))
use_rsfc_loadings_disc_top10 <- tail(use_rsfc_loadings_disc,floor(dim(use_rsfc_loadings_disc)[1]*.1)) #take top weighted 10%
use_rsfc_loadings_disc_top10 <- use_rsfc_loadings_disc_top10[order(use_rsfc_loadings_disc_top10[,1]),] #sort by ind
use_rsfc_loadings_disc <- use_rsfc_loadings_disc[order(use_rsfc_loadings_disc[,1]),] #sort by ind

rsfc_loadings_repl <- read.table(paste(workdir, '/replication/results/elasticnetCCA_scaled/discovery-selected-features/allreplication_rsFCloadings_v_CV1.csv', sep = ""))
rsfc_inds_repl <- read.table(paste(workdir, '/replication/results/elasticnetCCA_scaled/discovery-selected-features/allreplication_rsFCinds_v_CV1.csv', sep = ""))
use_rsfc_loadings_repl <- cbind(unlist(rsfc_inds_repl),unlist(rsfc_loadings_repl))
use_rsfc_loadings_repl_top10 <- tail(use_rsfc_loadings_repl,floor(dim(use_rsfc_loadings_repl)[1]*.1)) #take top weighted 10%
use_rsfc_loadings_repl_top10 <- use_rsfc_loadings_repl_top10[order(use_rsfc_loadings_repl_top10[,1]),] #sort by ind
use_rsfc_loadings_repl <- use_rsfc_loadings_repl[order(use_rsfc_loadings_repl[,1]),] #sort by ind

#find indices of rsFC in top 1% of both subsamples
overlap1 <- which(use_rsfc_loadings_disc_top10[,1] %in% use_rsfc_loadings_repl_top10[,1])
overlap2 <- which(use_rsfc_loadings_repl_top10[,1] %in% use_rsfc_loadings_disc_top10[,1])

overlap <- use_rsfc_loadings_disc_top10[overlap1, 1:2]
overlap <- cbind(overlap, use_rsfc_loadings_disc_top10[overlap2, 1:2])

#check that indices match
table(overlap[,1] == overlap[,3])

#check how many of the signs match
sum(sign(overlap[,2]) == sign(overlap[,4]))

#what percentage of the top 1% is this?
sum(sign(overlap[,2]) == sign(overlap[,4])) / dim(use_rsfc_loadings_disc_top10)[1]

#calculate correlation of disc and repl loadings
cor.test(use_rsfc_loadings_disc[,2], use_rsfc_loadings_repl[,2])
bothloadings <- cbind(use_rsfc_loadings_disc[,2], use_rsfc_loadings_repl[,2])
icc(bothloadings, model = "twoway", unit = "average", type = "consistency")


```
## Compare rsFC loadings across subsamples (Discovery versus Replication with Replication-selected features)
```{r}
rsfc_loadings_disc <- read.table(paste(workdir, '/discovery/results/elasticnetCCA_scaled/alldiscovery_rsFCloadings_v_CV1.csv', sep = ""))
rsfc_inds_disc <- read.table(paste(workdir, '/discovery/results/elasticnetCCA_scaled/alldiscovery_rsFCinds_v_CV1.csv', sep = ""))
use_rsfc_loadings_disc <- cbind(unlist(rsfc_inds_disc),unlist(rsfc_loadings_disc))
use_rsfc_loadings_disc <- tail(use_rsfc_loadings_disc,floor(dim(use_rsfc_loadings_disc)[1]*.1)) #take top weighted 10%
use_rsfc_loadings_disc <- use_rsfc_loadings_disc[order(use_rsfc_loadings_disc[,1]),]

rsfc_loadings_repl <- read.table(paste(workdir, '/replication/results/elasticnetCCA_scaled/replication-selected-features/allreplication_rsFCloadings_v_CV1.csv', sep = ""))
rsfc_inds_repl <- read.table(paste(workdir, '/replication/results/elasticnetCCA_scaled/replication-selected-features/allreplication_rsFCinds_v_CV1.csv', sep = ""))
use_rsfc_loadings_repl <- cbind(unlist(rsfc_inds_repl),unlist(rsfc_loadings_repl))
use_rsfc_loadings_repl <- tail(use_rsfc_loadings_repl,floor(dim(use_rsfc_loadings_repl)[1]*.1)) #take top weighted 10%
use_rsfc_loadings_repl <- use_rsfc_loadings_repl[order(use_rsfc_loadings_repl[,1]),]

#find indices of rsFC in top 1% of both subsamples
overlap1 <- which(use_rsfc_loadings_disc[,1] %in% use_rsfc_loadings_repl[,1])
overlap2 <- which(use_rsfc_loadings_repl[,1] %in% use_rsfc_loadings_disc[,1])

overlap <- use_rsfc_loadings_disc[overlap1, 1:2]
overlap <- cbind(overlap, use_rsfc_loadings_repl[overlap2, 1:2])

#check that indices match
table(overlap[,1] == overlap[,3])

#check how many of the signs match
sum(sign(overlap[,2]) == sign(overlap[,4]))

#what percentage of the top 1% is this?
sum(sign(overlap[,2]) == sign(overlap[,4])) / dim(use_rsfc_loadings_disc)[1]
```


# Exploratory analysis: enCCA with network-level rsFC features

## Figure 7: fitted and imputed canonical correlations across subsamples
```{r}
canoncorrsnetwork1 <- read.csv(paste(workdir, '/network-level/results/stricterexclusion/allsubsets_discoveryweights_canoncorrs1.csv', sep = ""), header = FALSE)
names(canoncorrsnetwork1) <- c("corr", "pval", "CI95lower", "CI95upper")
canoncorrsnetwork1$CV <- 1
canoncorrsnetwork1$Subsample <- c("Discovery", "Test", "Replication")

canoncorrsnetwork2 <- read.csv(paste(workdir, '/network-level/results/stricterexclusion/allsubsets_discoveryweights_canoncorrs2.csv', sep = ""), header = FALSE)
names(canoncorrsnetwork2) <- c("corr", "pval", "CI95lower", "CI95upper")
canoncorrsnetwork2$CV <- 2
canoncorrsnetwork2$Subsample <- c("Discovery", "Test", "Replication")

canoncorrsnetwork3 <- read.csv(paste(workdir, '/network-level/results/stricterexclusion/allsubsets_discoveryweights_canoncorrs3.csv', sep = ""), header = FALSE)
names(canoncorrsnetwork3) <- c("corr", "pval", "CI95lower", "CI95upper")
canoncorrsnetwork3$CV <- 3
canoncorrsnetwork3$Subsample <- c("Discovery", "Test", "Replication")

figdata <- rbind(canoncorrsnetwork1, canoncorrsnetwork2, canoncorrsnetwork3)
figdata[1,3] <- NA
figdata[1,4] <- NA
figdata[4,3] <- NA
figdata[4,4] <- NA
figdata[7,3] <- NA
figdata[7,4] <- NA

ggplot(figdata, aes(x=CV, y=corr, fill=Subsample, color=Subsample)) +
  geom_bar(stat="identity", position = "dodge", alpha=0.7) +
  scale_fill_manual(values = c("#48AEE1", "#EC6C53","#67C60F")) +
  scale_color_manual(values = c("#48AEE1", "#EC6C53","#67C60F")) +
  geom_errorbar( aes(x=CV, ymin=CI95lower, ymax=CI95upper), alpha=0.9, size=0.5, position = "dodge") +
  labs(x="Canonical variate pair (signature)", y="Canonical correlation") +
  theme_bw() + 
  theme(panel.grid.minor = element_blank(), panel.grid.major.x = element_blank()) +
  theme(legend.position = c(0.9, 0.85)) +
  theme(panel.border = element_blank(), axis.line = element_line()) +
  geom_hline(yintercept=0, linetype="dashed")


ggsave(paste(workdir, '/figures/fig7_network-level.png', sep = ""), width = 8, height = 6, dpi = 1000)


```


## Figure 8: Compare symptom weights across subsamples (Discovery versus Replication)
```{r}
signflip <- -1
netclin_loadings_disc <- signflip*read.table(paste(workdir, '/network-level/results/stricterexclusion/alldiscovery_clinloadings_u_CV1norm.csv', sep = ""))
netclin_loadings_repl <- signflip*read.table(paste(workdir, '/network-level/results/stricterexclusion/allreplication_clinloadings_u_CV1norm.csv', sep = ""))

cor.test(as.numeric(unlist(netclin_loadings_disc)), as.numeric(unlist(netclin_loadings_repl)))

icc(cbind(netclin_loadings_disc, netclin_loadings_repl), model = "twoway", unit = "average", type = "consistency")

clin_labels <- c("Obsessions", "SomethingBad", "Perfect", "Guilty", "Compulsions", "StrangeBehavior", "StrangeIdeas", "Worries")

#discovery
#clin_loadings <- netclin_loadings_disc
#replication
clin_loadings <- netclin_loadings_repl

names(clin_loadings) <- "Loadings"
clin_loadings$Symptoms <- factor(clin_labels, levels = rev(clin_labels), ordered = TRUE)
levels(clin_loadings$Symptoms)

if(sum(clin_loadings$Loadings>0) == 8) { #if all loadings are positive
  clin_visual <- ggplot(data=clin_loadings, aes(x=Symptoms, y=Loadings, fill=Loadings>0)) + 
  geom_bar(stat="identity") + 
  scale_fill_manual(values = c("#1ABA9B")) +

  ylim(-50,60) +
  labs(x = "", y = "Loading") +
  coord_flip() +
  theme_classic() +
  theme(axis.text = element_text(size = 17), axis.title = element_text(size = 17)) +
  theme (legend.position = "none")
} else {
  clin_visual <- ggplot(data=clin_loadings, aes(x=Symptoms, y=Loadings, fill=Loadings>0)) + 
    geom_bar(stat="identity") + 
    scale_fill_manual(values = c("#E36C3F", "#1ABA9B")) +
    ylim(-50,60) +
    labs(x = "", y = "Loading") +
    coord_flip() +
    theme_classic() +
    theme(axis.text = element_text(size = 17), axis.title = element_text(size = 17)) +
    theme (legend.position = "none")
}

clin_visual
ggsave(paste(workdir, '/figures/fig8-replication.png', sep = ""), width = 8, height = 6, dpi = 1000)

```

## Table 7: Compare network-level rsFC loadings across subsamples (Discovery versus Replication)
```{r}
netrsfc_loadings_disc <- signflip*read.table(paste(workdir, '/network-level/results/alldiscovery_rsFCloadings_v_CV1.csv', sep = ""))
netrsfc_inds_disc <- read.table(paste(workdir, '/network-level/results/alldiscovery_rsFCinds_v_CV1.csv', sep = ""))
use_netrsfc_loadings_disc <- cbind(unlist(netrsfc_inds_disc),unlist(netrsfc_loadings_disc))
all_rsfc_loadings_disc <- use_netrsfc_loadings_disc[order(use_netrsfc_loadings_disc[,1]),]


netrsfc_loadings_repl <- signflip*read.table(paste(workdir, '/network-level/results/allreplication_rsFCloadings_v_CV1.csv', sep = ""))
netrsfc_inds_repl <- read.table(paste(workdir, '/network-level/results/allreplication_rsFCinds_v_CV1.csv', sep = ""))
use_netrsfc_loadings_repl <- cbind(unlist(netrsfc_inds_repl),unlist(netrsfc_loadings_repl))
all_rsfc_loadings_repl <- use_netrsfc_loadings_repl[order(use_netrsfc_loadings_repl[,1]),]

cor.test(all_rsfc_loadings_disc[,2], all_rsfc_loadings_repl[,2])

icc(cbind(all_rsfc_loadings_disc[,2], all_rsfc_loadings_repl[,2]), model = "twoway", unit = "average", type = "consistency")

use_netrsfc_loadings_disc <- tail(use_netrsfc_loadings_disc,25) #take top weighted 25
use_netrsfc_loadings_disc <- apply(use_netrsfc_loadings_disc,2, rev) #reverse order
use_netrsfc_loadings_repl <- tail(use_netrsfc_loadings_repl,25) #take top weighted 25
use_netrsfc_loadings_repl <- apply(use_netrsfc_loadings_repl,2, rev) #reverse order

labels(connmats_raw_discovery)[[2]][use_netrsfc_loadings_disc[,1]]

labels(connmats_raw_replication)[[2]][use_netrsfc_loadings_repl[,1]]
```

