---
title: "Exploring rsfMRI Signatures of OCS Using Big Data and Machine Learning"
author: "Tracey Shi"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_float: no
  github_document:
    toc: yes
  html_notebook:
    df_print: paged
    highlight: tango
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_float: no
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
    fig_width: 3
    fig_height: 2.75
    fig_caption: true
---

# Setup

```{r}
workdir <- '/Users/marshlab2/Desktop/OCS_bigdata_ML_local'

```

# Data

Data is used from the Adolescent Brain Cognitive Development (ABCD) study.

## ABCD Data

ABCD release 4.0 on NDA (https://dx.doi.org/10.15154/1523041). Accessed 2022.01.13:

* abcd_betnet02.txt
* abcd_cbcl01.txt
* abcd_lt01.txt
* abcd_mri01.txt
* abcd_imgincl01.txt
* abcd_ksads01.txt

Baseline ABCD neuroimaging data preprocessed through DCAN-ABCD pipeline from DCAN-Lab, collection 3165 on NDA (https://nda.nih.gov/edit_collection.html?id=3165):

* Follow download instructions on Github page (https://github.com/DCAN-Labs/nda-abcd-s3-downloader) - use this Python application to download the actual connectivity matrices 
* participants.tsv (part of Collection Documents on NDA; accessed 2021.04.27)

## Inclusion criteria for discovery and test datasets (from ABCD study)

Eligible participants:

* abcd_imgincl01 file, "rsfmri include" == 1
* Have 10min pconn.nii file in data collection 3165 from DCANlab

Combined criteria for inclusion:

* Completed T1 and rsfmri scans, passed QA
* No significant anatomical findings
* At least 10min of motion-free resting state data after censoring out frames with FD>0.2mm

# Load and clean data and divide into subsets (using DCAN-Labs groups)

## Load data from various files

```{r}
# Load ABCD release 4.0 file abcd_imgincl01.txt (imaging QA inclusion/exclusion)
imgincl01 <- read.table(paste(workdir, '/data/abcd_imgincl01.txt', sep = ""), header = TRUE)
# drop first line of text (variable descriptions)
imgincl01 <- imgincl01[-1, ]

# Load ABCD release 4.0 file abcd_lt01.txt (age, sex)
lt01 <- read.table(paste(workdir, '/data/abcd_lt01.txt', sep = ""), header = TRUE)
# drop first line of text (variable descriptions)
lt01 <- lt01[-1, ]

# Load ABCD release 4.0 file abcd_mri01.txt (mri_info_deviceserialnumber)
mri01 <- read.table(paste(workdir, '/data/abcd_mri01.txt', sep = ""), header = TRUE)
# drop first line of text (variable descriptions)
mri01 <- mri01[-1, ]

## betnet02 (rsfMRI head motion (mean framewise displacement))
betnet02 <- read.table(paste(workdir, '/data/abcd_betnet02.txt', sep = ""), header = TRUE, sep = '\t')
# drop first line of text (variable descriptions)
betnet02 <- betnet02[-1, ]

# Load CBCL parent-reported questionnaire abcd_cbcl01.txt (used to calculate CBCL-OCS)
cbcl01 <- read.table(paste(workdir, '/data/abcd_cbcl01.txt', sep = ""), header = TRUE)

```

## Concatenate useful variables across ABCD files
```{r}
# Extract subject IDs for participants flagged for inclusion of baseline rsfMRI by ABCD QA process
abcd_incl <- imgincl01[imgincl01$eventname == 'baseline_year_1_arm_1', ]
paste('participants with imaging data in ABCD baseline data release: ', nrow(abcd_incl))
abcd_incl <- abcd_incl[abcd_incl$imgincl_rsfmri_include == 1, ]
paste('participants with baseline rsfMRI data marked for inclusion based on ABCD QA: ', nrow(abcd_incl))

# Remove non-baseline data from other ABCD files, keep only relevant variables
lt011y <- lt01[lt01$eventname == '1_year_follow_up_y_arm_1', c('subjectkey','interview_age')]
lt012y <- lt01[lt01$eventname == '2_year_follow_up_y_arm_1', c('subjectkey','interview_age')]
lt01 <- lt01[lt01$eventname == 'baseline_year_1_arm_1', c('subjectkey','interview_age','sex','site_id_l')]
mri01 <- mri01[mri01$eventname == 'baseline_year_1_arm_1', c('subjectkey','mri_info_deviceserialnumber')]
betnet02 <- betnet02[betnet02$eventname == 'baseline_year_1_arm_1', c('subjectkey','rsfmri_c_ngd_meanmotion')] #Average framewise displacement in mm

# double-check we are keeping the correct CBCL-OCS items
cbcl01[1, c(18,40,41,61,82,100,101,128)] #to view column descriptions
# drop first line of text (variable descriptions)
cbcl01 <- cbcl01[-1, ]
subjectkey <- cbcl01[cbcl01$eventname == 'baseline_year_1_arm_1', c('subjectkey')]
cbcl01 <- cbcl01[cbcl01$eventname == 'baseline_year_1_arm_1', c(18,40,41,61,82,100,101,128)]
#rename columns
names(cbcl01) <- c('obsessions','something_bad','perfect','guilty','compulsions','strange_behavior','strange_ideas','worries')
#convert columns to numeric from factor
cbcl01[ ,1:8] <- apply(cbcl01[ ,1:8], 2, function(x) as.numeric(as.character(x)))
cbcl01$cbcl_ocs_score <- rowSums(cbcl01)
cbcl01$subjectkey <- subjectkey

# Use imaging QA participant list as a template, merge in other variables
data_abcd <- abcd_incl[ , c('subjectkey', 'imgincl_rsfmri_include')]
data_abcd <- merge(data_abcd, lt01, by = 'subjectkey', all.x = TRUE)
data_abcd <- merge(data_abcd, mri01, by = 'subjectkey', all.x = TRUE)
data_abcd <- merge(data_abcd, betnet02, by = 'subjectkey', all.x = TRUE)
data_abcd <- merge(data_abcd, cbcl01, by = 'subjectkey', all.x = TRUE)

#visually check/confirm duplicates
duplicates1 <- subset(data_abcd,duplicated(data_abcd$subjectkey))
duplicates2 <- subset(data_abcd,duplicated(data_abcd$subjectkey, fromLast = TRUE))
#if needed, update both copies of the duplicate entry to have correct mri_info_deviceserialnumber
# for(d in 1:nrow(duplicates1)) {
#   goodinfo <- as.factor(paste(duplicates1$mri_info_deviceserialnumber[d], duplicates2$mri_info_deviceserialnumber[d], sep=''))
#   data_abcd$mri_info_deviceserialnumber[as.numeric(row.names(duplicates1)[d])] <- goodinfo
#   data_abcd$mri_info_deviceserialnumber[as.numeric(row.names(duplicates2)[d])] <- goodinfo
# }
# data_abcd <- data_abcd[!duplicated(data_abcd$subjectkey), ]
#rm(d)
rm(duplicates1, duplicates2)
```

Review variables for missingness
```{r}
# data release 4.0 no longer has missing values for mri_info_deviceserialnumber
nrow(data_abcd[data_abcd$mri_info_deviceserialnumber=='',])

#no other variables have missing values
nrow(data_abcd[data_abcd$imgincl_rsfmri_include=='',]) #can't be missing as participant list was defined off this
nrow(data_abcd[data_abcd$site=='',])
nrow(data_abcd[data_abcd$interview_age=='',])
nrow(data_abcd[data_abcd$sex=='',])

nrow(data_abcd[data_abcd$cbcl_ocs_score=='',]) #some of these are missing in data release 4.0 (5 missing)
table(data_abcd$cbcl_ocs_score)


```

## Format to combine with DCAN-Labs data

```{r}
#Reformat subjectkey to match DCAN participant ID formatting, add as new field 'participant_id'
participant_id <- as.character(data_abcd$subjectkey)
participant_id <- paste("sub-", participant_id, sep = "")
participant_id <- sub("_", "", participant_id)
participant_id <- as.data.frame(participant_id)
data_abcd <- cbind(data_abcd, participant_id)
rm(participant_id)

#Divide into discovery and test sets, format, and save as a subject list for DCAN downloader
#load participants.tsv from DCAN lab
participants_dcan <- read.table(paste('/Users/marshlab2/Dropbox/SimpsonMarshPhD/OCS_bigdata_ML','/scripts/nda-abcd-s3-downloader-master/Collection_Documents/participants_v1.0.0/participants.tsv', sep = ""), header = TRUE)

#keep only baseline data, participant ID and matched_group variables
participants_dcan <- participants_dcan[participants_dcan$session_id == 'ses-baselineYear1Arm1', c('participant_id', 'matched_group','race_ethnicity','handedness')]

subject_list_merged <- merge(data_abcd, participants_dcan, by = 'participant_id', all.x = TRUE)

# 44 participants in the DCAN spreadsheet have 2 entries (rows)
# each of these participants has 2 rows that are identical except for site (so each participant has 2 sites listed
# but none of these duplicates have contradictory matched_group assignments, so it doesn't affect this...
#duplicates1 <- subset(subject_list_merged,duplicated(subject_list_merged$participant_id))
#duplicates2 <- subset(subject_list_merged,duplicated(subject_list_merged$participant_id, fromLast = TRUE))
#duplicates <- rbind(duplicates1, duplicates2)
#rm(duplicates1, duplicates2, duplicates)
# remove duplicates
subject_list_merged <- subject_list_merged[!duplicated(subject_list_merged$participant_id), ]

subject_list_discovery <- subject_list_merged[subject_list_merged$matched_group == 1, ]
subject_list_discovery <- subject_list_discovery[!is.na(subject_list_discovery$participant_id), ] #remove rows with all NAs
#write.table(subject_list_discovery[ , c("participant_id")], paste(workdir, '/scripts/nda-abcd-s3-downloader-master/subject_list_discovery.txt', sep = ""), quote = FALSE, row.names = FALSE, col.names = FALSE)

subject_list_test <- subject_list_merged[subject_list_merged$matched_group == 3, ]
subject_list_test <- subject_list_test[!is.na(subject_list_test$participant_id), ] #remove rows with all NAs
#write.table(subject_list_test[ , c("participant_id")], paste(workdir, '/scripts/nda-abcd-s3-downloader-master/subject_list_test.txt', sep = ""), quote = FALSE, row.names = FALSE, col.names = FALSE)

subject_list_replication <- subject_list_merged[subject_list_merged$matched_group == 2, ]
subject_list_replication <- subject_list_replication[!is.na(subject_list_replication$participant_id), ] #remove rows with all NAs
#write.table(subject_list_replication[ , c("participant_id")], paste(workdir, '/scripts/nda-abcd-s3-downloader-master/subject_list_replication.txt', sep = ""), quote = FALSE, row.names = FALSE, col.names = FALSE)

# Clean up memory
rm(imgincl01, lt01, mri01, cbcl01, betnet02, participants_dcan, abcd_incl, subjectkey)

# Simple stats
table(subject_list_discovery$imgincl_rsfmri_incl)
table(subject_list_test$imgincl_rsfmri_incl)
table(subject_list_replication$imgincl_rsfmri_incl)

table(subject_list_discovery$sex)[1:2]
table(subject_list_test$sex)
chisq.test(subject_list_merged$matched_group, subject_list_merged$sex)

mean(as.numeric(as.character((subject_list_discovery$interview_age))), na.rm = TRUE)
sd(as.numeric(as.character((subject_list_discovery$interview_age))), na.rm = TRUE)
mean(as.numeric(as.character((subject_list_replication$interview_age))), na.rm = TRUE)
sd(as.numeric(as.character((subject_list_replication$interview_age))), na.rm = TRUE)
t.test(as.numeric(as.character((subject_list_discovery$interview_age))),as.numeric(as.character((subject_list_replication$interview_age))))

mean(as.numeric(as.character((subject_list_discovery$cbcl_ocs_score))), na.rm = TRUE)
sd(as.numeric(as.character((subject_list_discovery$cbcl_ocs_score))), na.rm = TRUE)
mean(as.numeric(as.character((subject_list_replication$cbcl_ocs_score))), na.rm = TRUE)
sd(as.numeric(as.character((subject_list_replication$cbcl_ocs_score))), na.rm = TRUE)
t.test(as.numeric(as.character((subject_list_discovery$cbcl_ocs_score))),as.numeric(as.character((subject_list_replication$cbcl_ocs_score))))

table(subject_list_discovery$race_ethnicity)
table(subject_list_replication$race_ethnicity)

#ocs_discovery <- hist(as.numeric(as.character((subject_list_discovery$cbcl_ocs_score))), main="Discovery", freq=FALSE, xlab="CBCL-OCS score", xlim=c(0,16))
#ocs_test <- hist(as.numeric(as.character((subject_list_test$cbcl_ocs_score))), main="Test", freq=FALSE, xlab="CBCL-OCS score", xlim=c(0,16))
#ocs_replication <- hist(as.numeric(as.character((subject_list_replication$cbcl_ocs_score))), main="Replication", freq=FALSE, xlab="CBCL-OCS score", xlim=c(0,16))

hist(as.numeric(as.character((subject_list_discovery$cbcl_ocs_score))), main="Discovery", freq=FALSE, xlab="CBCL-OCS Total Score", xlim=c(0,16), ylim=c(0,0.5), right = FALSE, layout=c(1,3))
hist(as.numeric(as.character((subject_list_test$cbcl_ocs_score))), main="Test", freq=FALSE, xlab="CBCL-OCS Total Score", xlim=c(0,16), ylim=c(0,0.5), right = FALSE)
hist(as.numeric(as.character((subject_list_replication$cbcl_ocs_score))), main="Replication", freq=FALSE, xlab="CBCL-OCS Total Score", xlim=c(0,16), ylim=c(0,0.5), right = FALSE)
```

## Download and format imaging data

1. Use DCAN-Labs' ABCD downloader (python3) to download 10minute pconn.nii files for these subject lists. DCAN-Labs downloader available here: https://github.com/DCAN-Labs/nda-abcd-s3-downloader

  # python3 download_focused.py -i datastructure_manifest.txt -o FILES_discovery -s subject_list_discovery.txt -l LOGS -d data_subsets_focused.txt
  # python3 download_focused.py -i datastructure_manifest.txt -o FILES_test -s subject_list_test.txt -l LOGS -d data_subsets_focused.txt

2. Download and install/add to path Matlab toolbox for working with cifti files: https://github.com/Washington-University/cifti-matlab

3. Remove duplicate raw connectivity features by taking upper or lower triangular matrix (without the diagonal)

4. Concatenate raw connectivity features for all participants in each subset (discovery, validation, replication)

5. Import raw connectivity features into R

```{r}
library(R.matlab)
connmats_raw_subset <- readMat('/Users/marshlab2/Desktop/OCS_bigdata_ML_local/data/dropbox_altstorage/cf_mat_raw_replication.mat')
connmats_raw_subset <- connmats_raw_subset[[1]]

```


## Extract clinical data and covariates (demographic, administrative, head motion)

Clinical data: 8 individual item scores from CBCL comprising the CBCL-OCS scale
Covariates: Age, sex, race, site
```{r}
#import (from Matlab) subject list denoting which participants have valid connmat
matlab_subject_list_subset <- read.table('/Users/marshlab2/Desktop/OCS_bigdata_ML_local/data/dropbox_altstorage/subj_list_raw_replication.txt', header = TRUE, sep = ",")
names(matlab_subject_list_subset) = c('participant_id','matlab_include')

#merge with covariates and clinical data file
data_subset <- merge(matlab_subject_list_subset, subject_list_replication, by = 'participant_id', all = TRUE)

#clinical data (CBCL-OCS individual items)
cbcl_ocs_subset <- data_subset[data_subset$matlab_include==1,c('obsessions','something_bad','perfect','guilty','compulsions','strange_behavior','strange_ideas','worries')]

#covariates
covariates_subset <- data_subset[data_subset$matlab_include==1,c('interview_age','sex','site_id_l','mri_info_deviceserialnumber','race_ethnicity','rsfmri_c_ngd_meanmotion')]

# #doublecheck that all IDs are in same order after merge
ids_subset <- data_subset[data_subset$matlab_include==1,c('participant_id','matlab_include')]
ids_subset$id_subjectlist <- matlab_subject_list_subset[matlab_subject_list_subset$matlab_include==1,c('participant_id')]
ids_subset$check <- (ids_subset$participant_id == ids_subset$id_subjectlist)
table(ids_subset$check)

subjectkey_subset <- data_subset[data_subset$matlab_include==1,c('participant_id','subjectkey')]

#check/reformat all variables to appropriate class
#age (numeric)
covariates_subset$interview_age <- as.numeric(as.character(covariates_subset$interview_age))
covariates_subset$scaled_age <- scale(covariates_subset$interview_age) #standardize variable
mean(covariates_subset$interview_age)
sd(covariates_subset$interview_age)
class(covariates_subset$scaled_age)

#sex (factor)
class(covariates_subset$sex)
#site (factor)
class(covariates_subset$site_id_l)
levels(covariates_subset$site_id_l)
covariates_subset$site_id_l <- droplevels(covariates_subset$site_id_l) #drop unused levels
levels(covariates_subset$site_id_l)
#scanner (factor)
class(covariates_subset$mri_info_deviceserialnumber)
table(covariates_subset$mri_info_deviceserialnumber)
covariates_subset$mri_info_deviceserialnumber <- droplevels(covariates_subset$mri_info_deviceserialnumber)
levels(covariates_subset$mri_info_deviceserialnumber)
#race/ethnicity (factor)
class(covariates_subset$race_ethnicity) #defaults to integer
covariates_subset$race_ethnicity <- as.factor(covariates_subset$race_ethnicity)
class(covariates_subset$race_ethnicity)
levels(covariates_subset$race_ethnicity)
table(covariates_subset$race_ethnicity)
# remove participants coded as 888 as causes issue with covariate regression from clinical features
# this affects 2 participants in discovery subsample, 6 in replication subsample
covariates_subset$race_ethnicity[covariates_subset$race_ethnicity=='888'] <- NA
covariates_subset$race_ethnicity <- droplevels(covariates_subset$race_ethnicity)
missingraceeth <- which(is.na(covariates_subset$race_ethnicity))
levels(covariates_subset$race_ethnicity)

#mean motion (numeric)
covariates_subset$rsfmri_c_ngd_meanmotion <- as.numeric(as.character(covariates_subset$rsfmri_c_ngd_meanmotion))
covariates_subset$scaled_motion <- scale(covariates_subset$rsfmri_c_ngd_meanmotion) #standardize variable
mean(covariates_subset$rsfmri_c_ngd_meanmotion, na.rm = TRUE)
sd(covariates_subset$rsfmri_c_ngd_meanmotion, na.rm = TRUE)
class(covariates_subset$scaled_motion)

rm(subject_list_merged)
rm(subject_list_discovery)
#rm(subject_list_replication)
rm(subject_list_test)
```

```{r}
#drop the participants without race/ethnicity data
#NB: this is done manually here because setting 'na.action = na.exclude' doesn't work for zeroinfl function
# DON'T RUN THIS BLOCK IF missingraceeth is empty
cbcl_ocs_subset <- cbcl_ocs_subset[c(-missingraceeth),]
covariates_subset <- covariates_subset[c(-missingraceeth),]
connmats_raw_subset <- connmats_raw_subset[c(-missingraceeth),]
ids_subset <- ids_subset[c(-missingraceeth),]
ids_subset <- ids_subset$participant_id
subjectkey_subset <- subjectkey_subset[c(-missingraceeth),]

# check for NA values
any(is.na(cbcl_ocs_subset))
any(is.na(connmats_raw_subset))
any(is.na(covariates_subset))
```

## Save raw clinical, rs-FC, and covariates data
```{r eval=FALSE, include=FALSE}
rm(data_abcd, data_discovery, matlab_subject_list_subset, missingraceeth, workdir)
save.image("~/Desktop/OCS_bigdata_ML_local/FINAL/replication/data/Data_prep/replication_raw_data.RData")

#save discovery subject ID list
write.csv(subjectkey_subset,file="~/Desktop/OCS_bigdata_ML_local/FINAL/replication/data/Data_prep/subjectkey_replication.csv",row.names=F)
```

## Regress out covariates and visualize rsFC features
```{r}
rsFC_features <- array(numeric(), dim(connmats_raw_subset)) #initialize array for rsFC features (after covariates are regressed out)
rsFC_covariate_coeffs <- array(numeric(), c(36, ncol(connmats_raw_subset)))
rsFC_feat_normaltest <- array(numeric(), c(2, ncol(connmats_raw_subset)))

for(i in 1:ncol(connmats_raw_subset)) {
  model_i <- lm(connmats_raw_subset[, i] ~ covariates_subset$scaled_age + covariates_subset$sex + covariates_subset$mri_info_deviceserialnumber + covariates_subset$race_ethnicity + covariates_subset$scaled_motion, na.action = na.exclude)
  rsFC_features[, i] <- residuals(model_i) #pads residuals with NA where predictor vars have NA values
  rsFC_covariate_coeffs[, i] <- model_i$coefficients
 
  #Shapiro-Wilk test for normality of residuals (i.e. the rsFC feature after covariate regression)
  # p>.05 means data are not significantly different from normal distribution
  shapiro_result <- shapiro.test(rsFC_features[, i])
  rsFC_feat_normaltest[1, i] <- shapiro_result$statistic
  rsFC_feat_normaltest[2, i] <- shapiro_result$p.value
}

names(rsFC_feat_normaltest) <- c('statistic','pval')
hist(rsFC_feat_normaltest[2, ], main = paste('Histogram of Shapiro-Wilk test p-values for each rsFC feature'), xlab = 'p-value')

#standardize (mean center, scale by SD)
rsFC_features_scaled <- scale(rsFC_features, center = TRUE, scale = TRUE)
write.csv(rsFC_features_scaled, '/Users/marshlab2/Desktop/OCS_bigdata_ML_local/FINAL/replication/data/Data_prep/rsFC_features_scaled.csv', row.names = FALSE, col.names = FALSE)

#write.table(rsFC_features, paste(workdir, '/data/ML_features/rsFC_features_discovery.txt', sep = ""), col.names=NA)
#write.table(rsFC_covariate_coeffs, paste(workdir, '/data/ML_features/rsFC_covariate_coeffs_discovery.txt', sep = ""), col.names=NA)
#write.table(rsFC_feat_normaltest, paste(workdir, '/data/ML_features/rsFC_feat_normaltest_discovery.txt', sep = ""), col.names=NA)

any(is.na(rsFC_features))


```

## Regress out covariates and visualize clinical features
```{r}
library(pscl)

class(cbcl_ocs_subset$obsessions)

# view cross tabulated data
table(cbcl_ocs_subset[,1], covariates_subset$race_ethnicity)
#table(cbcl_ocs_discovery[,1], covariates_discovery$site_id_l)
  
clin_features <- array(numeric(), dim(cbcl_ocs_subset)) #initialize array for rsFC features (after covariates are regressed out)
clin_covariate_coeffs_count <- array(numeric(), c(7, ncol(cbcl_ocs_subset))) #coefficients for poisson portion of model
clin_covariate_coeffs_zero <- array(numeric(), c(7, ncol(cbcl_ocs_subset))) #coefficients for binary portion of model
clin_feat_normaltest <- array(numeric(), c(2, ncol(cbcl_ocs_subset)))
  
for(i in 1:ncol(cbcl_ocs_subset)) {
  model_i <- zeroinfl(cbcl_ocs_subset[, i] ~ covariates_subset$scaled_age + covariates_subset$sex + covariates_subset$race_ethnicity, dist = "poisson", na.action = na.exclude)
  #model_i <- polr(cbcl_ocs_discovery[, i] ~ covariates_discovery$scaled_age + covariates_discovery$sex + covariates_discovery$race_ethnicity + covariates_discovery$site_id_l)
  clin_features[, i] <- model_i$residuals
  clin_covariate_coeffs_count[, i] <- model_i$coefficients$count
  clin_covariate_coeffs_zero[, i] <- model_i$coefficients$zero
 
  #Shapiro-Wilk test for normality of residuals (i.e. the rsFC feature after covariate regression)
  # p>.05 means data are not significantly different from normal distribution
  shapiro_result <- shapiro.test(clin_features[, i])
  clin_feat_normaltest[1, i] <- shapiro_result$statistic
  clin_feat_normaltest[2, i] <- shapiro_result$p.value
}

names(clin_feat_normaltest) <- c('statistic','pval')

#visualize - note that the clinical features after covariate regression are nowhere close to normally distributed
hist(clin_features[, 1])
hist(clin_features[, 2])
hist(clin_features[, 3])
hist(clin_features[, 4])
hist(clin_features[, 5])
hist(clin_features[, 6])
hist(clin_features[, 7])
hist(clin_features[, 8])

#standardize (mean center, scale by SD)
clin_features_scaled <- scale(clin_features, center = TRUE, scale = TRUE)
write.csv(clin_features_scaled, '/Users/marshlab2/Desktop/OCS_bigdata_ML_local/FINAL/replication/data/Data_prep/clin_features_scaled.csv', row.names = FALSE, col.names = FALSE)


#write.table(clin_features, paste(workdir, '/data/ML_features/clin_features_discovery.txt', sep = ""), col.names=NA)
#write.table(clin_covariate_coeffs_count, paste(workdir, '/data/ML_features/clin_covariate_coeffs_count_discovery.txt', sep = ""), col.names=NA)
#write.table(clin_covariate_coeffs_zero, paste(workdir, '/data/ML_features/clin_covariate_coeffs_zero_discovery.txt', sep = ""), col.names=NA)
#write.table(clin_feat_normaltest, paste(workdir, '/data/ML_features/clin_feat_normaltest_discovery.txt', sep = ""), col.names=NA)

any(is.na(clin_features))


```

Clean up workspace
```{r}
rm(clin_covariate_coeffs_count, clin_covariate_coeffs_zero, clin_feat_normaltest)
rm(connmats_raw_discovery, rsFC_covariate_coeffs, rsFC_feat_normaltest)
rm(covariates_discovery, data_abcd, data_discovery, matlab_subject_list_discovery, model_i, shapiro_result)
rm(d, goodinfo, i, missingraceeth)
```

## Partition data for cross-validation
```{r}
set.seed(12345) # set seed for reproducibility

# set multiple k-folds cross-validation parameters
kfolds_k <- 10
kfolds_times <- 10

paste("number of k-folds repetitions: ", kfolds_times) 
paste("number of cross-validation folds per rep: ", kfolds_k)

library(caret)

partitions <- createMultiFolds((1:dim(clin_features_scaled)[1]), k = kfolds_k, times = kfolds_times)
# partitions contains row indices of participants for each cross-validation-training set

set.seed(111)
partitions2 <- createMultiFolds((1:dim(clin_features_scaled)[1]), k = kfolds_k, times = 1)
```

Store indices of cross-validation partitions for Matlab/python
```{r eval=FALSE, include=FALSE}
#cross-validation partitions for enCCA
write.csv(partitions2$Fold01.Rep1, '/Users/marshlab2/Desktop/OCS_bigdata_ML_local/FINAL/replication/data/partitions2_rep01fold1.csv', row.names = FALSE, col.names = FALSE)
write.csv(partitions2$Fold02.Rep1, '/Users/marshlab2/Desktop/OCS_bigdata_ML_local/FINAL/replication/data/partitions2_rep01fold2.csv', row.names = FALSE, col.names = FALSE)
write.csv(partitions2$Fold03.Rep1, '/Users/marshlab2/Desktop/OCS_bigdata_ML_local/FINAL/replication/data/partitions2_rep01fold3.csv', row.names = FALSE, col.names = FALSE)
write.csv(partitions2$Fold04.Rep1, '/Users/marshlab2/Desktop/OCS_bigdata_ML_local/FINAL/replication/data/partitions2_rep01fold4.csv', row.names = FALSE, col.names = FALSE)
write.csv(partitions2$Fold05.Rep1, '/Users/marshlab2/Desktop/OCS_bigdata_ML_local/FINAL/replication/data/partitions2_rep01fold5.csv', row.names = FALSE, col.names = FALSE)
write.csv(partitions2$Fold06.Rep1, '/Users/marshlab2/Desktop/OCS_bigdata_ML_local/FINAL/replication/data/partitions2_rep01fold6.csv', row.names = FALSE, col.names = FALSE)
write.csv(partitions2$Fold07.Rep1, '/Users/marshlab2/Desktop/OCS_bigdata_ML_local/FINAL/replication/data/partitions2_rep01fold7.csv', row.names = FALSE, col.names = FALSE)
write.csv(partitions2$Fold08.Rep1, '/Users/marshlab2/Desktop/OCS_bigdata_ML_local/FINAL/replication/data/partitions2_rep01fold8.csv', row.names = FALSE, col.names = FALSE)
write.csv(partitions2$Fold09.Rep1, '/Users/marshlab2/Desktop/OCS_bigdata_ML_local/FINAL/replication/data/partitions2_rep01fold9.csv', row.names = FALSE, col.names = FALSE)
write.csv(partitions2$Fold10.Rep1, '/Users/marshlab2/Desktop/OCS_bigdata_ML_local/FINAL/replication/data/partitions2_rep01fold10.csv', row.names = FALSE, col.names = FALSE)
```

```{r eval=FALSE, include=FALSE}
#cross-validation partitions for feature selection
write.csv(partitions$Fold01.Rep10, '/Users/marshlab2/Desktop/OCS_bigdata_ML_local/FINAL/replication/data/Data_prep/PartitionsFeatureSelection/partitions_rep10fold1.csv', row.names = FALSE, col.names = FALSE)
write.csv(partitions$Fold02.Rep10, '/Users/marshlab2/Desktop/OCS_bigdata_ML_local/FINAL/replication/data/Data_prep/PartitionsFeatureSelection/partitions_rep10fold2.csv', row.names = FALSE, col.names = FALSE)
write.csv(partitions$Fold03.Rep10, '/Users/marshlab2/Desktop/OCS_bigdata_ML_local/FINAL/replication/data/Data_prep/PartitionsFeatureSelection/partitions_rep10fold3.csv', row.names = FALSE, col.names = FALSE)
write.csv(partitions$Fold04.Rep10, '/Users/marshlab2/Desktop/OCS_bigdata_ML_local/FINAL/replication/data/Data_prep/PartitionsFeatureSelection/partitions_rep10fold4.csv', row.names = FALSE, col.names = FALSE)
write.csv(partitions$Fold05.Rep10, '/Users/marshlab2/Desktop/OCS_bigdata_ML_local/FINAL/replication/data/Data_prep/PartitionsFeatureSelection/partitions_rep10fold5.csv', row.names = FALSE, col.names = FALSE)
write.csv(partitions$Fold06.Rep10, '/Users/marshlab2/Desktop/OCS_bigdata_ML_local/FINAL/replication/data/Data_prep/PartitionsFeatureSelection/partitions_rep10fold6.csv', row.names = FALSE, col.names = FALSE)
write.csv(partitions$Fold07.Rep10, '/Users/marshlab2/Desktop/OCS_bigdata_ML_local/FINAL/replication/data/Data_prep/PartitionsFeatureSelection/partitions_rep10fold7.csv', row.names = FALSE, col.names = FALSE)
write.csv(partitions$Fold08.Rep10, '/Users/marshlab2/Desktop/OCS_bigdata_ML_local/FINAL/replication/data/Data_prep/PartitionsFeatureSelection/partitions_rep10fold8.csv', row.names = FALSE, col.names = FALSE)
write.csv(partitions$Fold09.Rep10, '/Users/marshlab2/Desktop/OCS_bigdata_ML_local/FINAL/replication/data/Data_prep/PartitionsFeatureSelection/partitions_rep10fold9.csv', row.names = FALSE, col.names = FALSE)
write.csv(partitions$Fold10.Rep10, '/Users/marshlab2/Desktop/OCS_bigdata_ML_local/FINAL/replication/data/Data_prep/PartitionsFeatureSelection/partitions_rep10fold10.csv', row.names = FALSE, col.names = FALSE)
```


Save R workspace for continued work in R

Do feature selection on unscaled features, but use scaled features for running CCA for cross-validation/hyperparam tuning!






Summary statistics by subsample
```{r}
mean(covariates_subset$interview_age)/12
sd(covariates_subset$interview_age)/12
range(covariates_subset$interview_age)/12

table(covariates_subset$sex)[["F"]]/length(covariates_subset$sex)

#CBCL-OCS total score
t <- rowSums(cbcl_ocs_subset)
mean(t)
sd(t)
hist(t,main = "Discovery", xlab = "CBCL-OCS Total Score", freq = FALSE, right = FALSE, ylim = range(0,0.5))
rm(t)

cbcl_ocs_subset_binary <- as.factor(cbcl_ocs_subset>0)
cbcl_ocs_subset_binary <- matrix(cbcl_ocs_subset_binary, nrow = dim(cbcl_ocs_subset)[1])

#individual CBCL-OCS items in-sample prevalence
table(cbcl_ocs_subset_binary[,1])[["TRUE"]]/length(cbcl_ocs_subset_binary[,1])
table(cbcl_ocs_subset_binary[,2])[["TRUE"]]/length(cbcl_ocs_subset_binary[,1])
table(cbcl_ocs_subset_binary[,3])[["TRUE"]]/length(cbcl_ocs_subset_binary[,1])
table(cbcl_ocs_subset_binary[,4])[["TRUE"]]/length(cbcl_ocs_subset_binary[,1])
table(cbcl_ocs_subset_binary[,5])[["TRUE"]]/length(cbcl_ocs_subset_binary[,1])
table(cbcl_ocs_subset_binary[,6])[["TRUE"]]/length(cbcl_ocs_subset_binary[,1])
table(cbcl_ocs_subset_binary[,7])[["TRUE"]]/length(cbcl_ocs_subset_binary[,1])
table(cbcl_ocs_subset_binary[,8])[["TRUE"]]/length(cbcl_ocs_subset_binary[,1])

```







